Regression Cube Analysis of Cohort Study Data

### 2015-03-30 draft.tex
added Steffis revisions

%% indicate IEEE Member or Student Member in form indicated below
\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernard Preim}
\authorfooter{
%% insert punctuation at end of each item
\item
 Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,niemann,preim\}@ovgu.de
\item
 Katrin Hegenscheid, Henry V{\"o}lzke are with Ernst-Moritz-Arndt University Greifswald, Germany. E-mail: \{katrin.hegenscheid,voelzke\}@uni-greifswald.de
}

@@

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
%%
A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
%%
A qualitative analysis was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
%%
%\com{Implications}.
%%
With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

\maketitle
%%
Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
%%
Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
%%
Furthermore, the insights can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
%%
During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then statistically analyzed.

An important epidemiological tool for deriving such features is the evaluation of a \emph{cohort study}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
%%
To reduce any selection bias, subjects are randomly invited without a focus on a specific disease.
%%
Hence, a wide range of features is acquired.
%
Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure, are gathered.
%%
Medical image data from non-radiating modalities, e.g., magnetic resonance imaging (MRI), is also acquired in modern studies.
%%
These data may be quantified based on user-defined landmarks, which describe attributes like the shape, the volume or the diameter of a structure.
%%

@@

In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to deriving insight and even new hypotheses through the data rather than using it just for confirmatory analysis.
%An alternative data-driven approach is described as follows.
%%
% \paragraph{Data Driven Hypothesis-Generation-based Analysis.}
% We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
% %%
% The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
% %%
% Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
%We focus on cohort studies are a major epidemiological tool for gathering epidemiological features.
We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
%%
%\paragraph{Data Types.}
%%
Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
%The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
%%
%The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

Medical image data is also acquired and analyzed in modern cohort studies.
%%
We incorporate image-derived features, but do not put focus on analyzing image data.
%%
More discussion of cohort study data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.
%%
%\paragraph{Image Data.}
% \com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
% %%
% Since the Rotterdam study, many modern cohort studies include medical image data.
% %%
% For ethical reasons, the imaging modalities usually not include ionizing radiation.
% %%
% The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
% %%
% These data are hard to analyze as they require segmentations to outline structures of interest.
% %%
% This process is prone to inter- and intra-observer variability when carried out manually.
% %%
% Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
% %%
% Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
\\\\
%\paragraph{Confounding Features.}
Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%Confounding features have to be considered in epidemiological analysis, as they influence both the exposure as well as the outcome of an analysis.
%%
%Features influencing the exposure as well as the outcome of an analysis are called confounders.
%%
The analysis model has to be adjusted by normalizing all included features towards the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely with increasing age.
%%
It also influences the general body condition and thereby almost all features acquired through cohort studies.
%%
Another important confounder is \emph{gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data.
%%
The presented work is based on regression analysis as well.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values describe the resulting function (the \emph{model}).
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value is in the range [0,1], where 1 encodes a perfect fit.

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
%Dependent \sim \sum_{i=1}^n Independent_i
\label{eq:RegressionNotation}
\end{equation}
%%
A example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
The most used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable,
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The type of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following:
%%
\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of a continuous type.
%%

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models as opposed to linear models do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behave different than $R^2$ values extracted from linear regression model.
%%
Comparisons have to be handled with care.

@@

\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study. % ("\emph{method loyality}").
%%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
%
%They slightly differ in clinical focus, acquired data and epidemiological research questions.
%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%
%Data are collected without focus on a group of diseases.
%
%This allows to query the data regarding many diseases and conditions.
%
%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data is too often only analyzed using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
%%
\\\\
The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} are closest related to ours regarding different aspects.
%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features towards the target feature using a plot matrix displaying models as contours.
%%
They also incorporate 3D visualizations for each pairwise combination, but mainly because its popularity with their target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
The main difference is their focus on analyzing one complex model in detail, yielding extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
%%
They also focus on metric features, while we process categorical data as well.
%%
%We share the matrix-arrangement of results depicting relationships between independent features in a model.

%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term Regression Cube for a comparative 3D representation of multiple regression models.
%% einheitlich 2D-XXX oder 2D XXX, meist ist es 2D XXX
Chan et al. \cite{Chan} use the same term to describe an extension of the 2D Scatterplot representation of a linear regression models (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to \cite{Piringer}, they focus on highlighting details of the included models, rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%

@@

\paragraph{Visual Analysis in Public Health.}
%%--
Shneiderman et al. \cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
%%
%They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
%%
%We focus on the latter.
%%
%Shneiderman and colleagues also present challenges, 
The presented challenges in this application comprise the systems to be sufficiently comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
%%
Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%
%%

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
% TODO: Do we need a preprocessing section?
%\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicate the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
%%
% unified -> singular [Steffi?]
A regression model is based on expert knowledge, there is no singular rule how to apply them to a given set of features. Thus, they have to be applied with care.
%%
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into regression analysis by the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

The major advantage of this notation is that it comes natural to anyone familiar with regression analysis since it is the standard way of expression.

@@

	\item $Z \sim X + Y$ calculates all combinations of two features towards all possible target features.

@@

When each regression takes about 50~ms calculation, we have to roughly wait 14~h for the calculation to complete.

@@

In general, many of them exhibit low or none correlation at all towards the target feature.

@@

The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.

@@

Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.

@@

\textbf{Show the Important} [\textbf{2.SI}]. The 3D-cube visualization acts as an overview over the whole data set.
%%
Regression models with large $R^2$ values can be spotted here fast, steering the users attention to the respective slice.

@@

The focus lies on creating an overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).

@@

Web technology is by design based on a client-server architecture, making it easy to outsource computationally heavy tasks on server-clusters and transferring results to the client device.

@@

\paragraph{File-Upload and Classification [1.AF].}

>>>>

%% indicate IEEE Member or Student Member in form indicated below
\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernhard Preim}
\authorfooter{
%% insert punctuation at end of each item
\item
 Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,niemann,preim\}@ovgu.de
\item
 Katrin Hegenscheid, Henry V{\"o}lzke are with Ernst-Moritz-Arndt University Greifswald, Germany. E-mail: \{katrin.hegenscheid,voelzke\}@uni-greifswald.de
}

@@

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization that reveals insights into the data by highlighting relationships.
%%
A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
%%
A qualitative analysis was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
%%
%\com{Implications}.
%%
With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

\maketitle
%%
Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
%%
Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
%%
Furthermore, the insights can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
%%
During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then are statistically analyzed.

An important epidemiological tool for deriving such features is the evaluation of a \emph{cohort study}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
%%
To reduce any selection bias, subjects are randomly invited without a focus on a specific disease.
%%
Hence, a wide range of features is acquired.
%
Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure, are gathered.
%%
Medical image data from non-radiating modalities, e.g., magnetic resonance imaging (MRI), is also acquired in modern studies.
%%
These data may be quantified based on user-defined landmarks, which describe attributes like the shape, volume or diameter of a structure.
%%

@@

In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%An alternative data-driven approach is described as follows.
%%
% \paragraph{Data Driven Hypothesis-Generation-based Analysis.}
% We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
% %%
% The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
% %%
% Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
%We focus on cohort studies are a major epidemiological tool for gathering epidemiological features.
We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
%%
%\paragraph{Data Types.}
%%
Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
%The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
%%
%The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

Medical image data is also acquired and analyzed in modern cohort studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of cohort study data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.
%%
%\paragraph{Image Data.}
% \com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
% %%
% Since the Rotterdam study, many modern cohort studies include medical image data.
% %%
% For ethical reasons, the imaging modalities usually not include ionizing radiation.
% %%
% The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
% %%
% These data are hard to analyze as they require segmentations to outline structures of interest.
% %%
% This process is prone to inter- and intra-observer variability when carried out manually.
% %%
% Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
% %%
% Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
\\\\
%\paragraph{Confounding Features.}
Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%Confounding features have to be considered in epidemiological analysis, as they influence both the exposure as well as the outcome of an analysis.
%%
%Features influencing the exposure as well as the outcome of an analysis are called confounders.
%%
The analysis model has to be adjusted by normalizing all included features towards the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely with increasing age.
%%
It also influences the general body condition and thereby almost all features acquired through cohort studies.
%%
Another important confounder is \emph{gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data.
%%
The presented work is based on regression analysis as well.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value is in the range [0,1], where 1 encodes a perfect fit.

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
%Dependent \sim \sum_{i=1}^n Independent_i
\label{eq:RegressionNotation}
\end{equation}
%%
An example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable,
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The type of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following:
%%
\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of a continuous type.
%%

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.

@@

\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study. % ("\emph{method loyality}").
%%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
%
%They slightly differ in clinical focus, acquired data and epidemiological research questions.
%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%
%Data are collected without focus on a group of diseases.
%
%This allows to query the data regarding many diseases and conditions.
%
%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data is too often only analyzed using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
%%
\\\\
The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} are closest related to ours regarding different aspects.
%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features towards the target feature using a plot matrix displaying models as contours.
%%
They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
The main difference is their focus on analyzing one complex model in detail, yielding extensive plots, while we analyze a large amount of simpler models in terms of \emph{different} features.
%%
They also focus on metric features, while we process categorical data as well.
%%
%We share the matrix-arrangement of results depicting relationships between independent features in a model.

%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
We use the term Regression Cube for a comparative 3D representation of multiple regression models.
%% einheitlich 2D-XXX oder 2D XXX, meist ist es 2D XXX
Chan et al. \cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to \cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%

@@

\paragraph{Visual Analysis in Public Health.}
%%--
Shneiderman et al. \cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
%%
%They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
%%
%We focus on the latter.
%%
%Shneiderman and colleagues also present challenges, 
The presented challenges in this application comprise the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
%%
Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%
%%

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
\begin{enumerate}
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
% TODO: Do we need a preprocessing section?
%\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicate the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
%%
% unified -> singular [Steffi?]
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into regression analysis by the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, since it is the standard way of expression.

@@

	\item $Z \sim X + Y$ calculates all combinations of two features w.r.t. all possible target features.

@@

When each regression takes about 50~ms calculation, we roughly have to wait 14~h for the calculation to complete.

@@

In general, many of them exhibit a low or no correlation at all in view of the target feature.

@@

The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$, which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.

@@

Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.

@@

\textbf{Show the Important} [\textbf{2.SI}]. The 3D cube visualization acts as an overview over the whole data set.
%%
Here, regression models with large $R^2$ values can be spotted fast, steering the user's attention to the respective slice.

@@

The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.

@@

By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.

@@

\paragraph{File Upload and Classification [1.AF].}
